Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-13 manuscript.tex
Update figures 2 and 5 and section 3.7

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by ultra-fast temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC3-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSummary, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

@@

\textbf{Dimensionality reduction.} The featurized trajectories were then projected into a low-dimensional space in preparation for microstate clustering. The standard approach to doing so is to employ time-lagged independent components analysis (tICA) to learn a linear projection into a low-dimensional embedding that maximally preserves the kinetic variance in the data.~\citep{Perez-Hernandez2013IdentificationConstruction, Schwantes2013ImprovementsNTL9, Noe2015KineticSimulation} In this work, we instead employ state-free reversible VAMPnets (SRVs) that can be conceived of as a nonlinear version of tICA. \citep{Chen} SRVs employ neural networks to learn flexible nonlinear functions of the trajectory featurization that better approximate the slow dynamical modes of the system and have been shown to produce substantially higher resolution MSMs than those developed using tICA in applications to WW-domain and Trp-cage mini-proteins.\citep{Chen, Sidky} Importantly, the SRV modes were learned independently for each system to best approximate the slow collective modes for that particular DNA sequence. SRVs were trained using the SRV package we previously developed (\url{https://github.com/hsidky/srv}) employing the default network architecture of two hidden layers each comprising 100 neurons and tanh activation functions, a learning rate of 0.001, and a batch size of 50,000. We adopted a lag time of $\tau$ = 1.2 ns as appropriate short to resolve the dynamical details of the hybridization/dehybridization dynamics. \citep{Phys2018MarkovValidation} As observed by Husic and Pande, it is conceptually incorrect to treat the lag time as a model hyperparameter to be self-consistently optimized via the VAMP-2 score, but rather as a physically-motivated choice designed to expose the dynamical motions relevant at a particular time and length scale of interest. \citep{Husic2017Note:Selection} As we shall show, this choice of lag time leads to converged and Markovian macrostate MSMs. We guarded against overfitting using five-fold cross-validation in which we divided the trajectory data into 50:50 training:validation splits. We observed plateau of the validation loss and no evidence of overfitting after 20 epochs of training for each system that required approximately 22 GPU-minutes on a single NVIDIA Tesla K80 GPU card. A VAMP-2 scoring of the cumulative kinetic variance explained as a function of number of SRV collective modes retained also showed no evidence of overfitting -- as would be evinced by separation of the training and validation VAMP-2 scores \citep{Sidky} -- and exhibited a knee for each of the four DNA sequences after the fifth, fourth, third, and second slow modes for AT-all, GC-end, GC-core, and GC-mix, respectively (\blauw{Fig.~S1} in the \blauw{Supporting Information}), and motivating the construction of 5D, 4D, 3D, and 3D embeddings, respectively.

@@

\textbf{Macrostate clustering.} The 200 microstates comprising each system were finally coarsened into our terminal macrostate MSM. The microstate transition matrix for each system was computed at a range of lag times $\tau$ and then diagonalized to recover the corresponding eigenvalues $\lambda_i$ and associated implied time scales $t_i = -\tau / \ln \left| \lambda_i \right|$.~\citep{Wehmeyer2019IntroductionV1.0} The implied time scale plots for the four DNA sequences are presented in \blauw{Fig.~S2}. We observe rapid convergence of the implied time scales $t_i$ with lag time $\tau$ for all systems, motivating the construction of high resolution macrostate MSMs at a lag time $\tau$ = 1.2 ns. For this choice of lag time, we recover 5, 4, 2, and 1 implied time scales for the AT-all, GC-end, GC-core, and GC-mid systems, respectively. The identification of $(i-1)$ implied time scales implies the presence of $(i-1)$ slow modes and motivates the coarsening of the system into $i$ macrostates. We estimate these $i$ macrostates by applying PCCA+ spectral clustering to the leading $(i-1)$ eigenvectors of the microstate transition matrix.~\citep{Roblitz2013FuzzyClassification, Weber2018ImplicationsSimulation, Kube2007AConformations} We then estimate the corresponding 6, 5, 3, and 2 macrostate transition matrices $\mathbf{P}$ for the AT-all, GC-end, GC-core, and GC-mid systems, respectively, by projecting the MD simulation trajectories into these discrete macrostates. These macrostate MSMs constitute our terminal kinetic models. We validate the Markovian nature of the four MSMs by subjecting them to the Chapman-Kolmogorov (CK) test \citep{Noe2009ConstructingSimulations,Phys2011MarkovValidation,  Wehmeyer2019IntroductionV1.0}. This test asserts that the transition matrix for a Markovian (i.e., memoryless) MSM constructed at a lag time $\tau$ should satisfy the condition $\mathbf{P}(k \tau) = \mathbf{P}^k(\tau)$, which states that $k$ successive applications of the transition matrix constructed at a lag time $\tau$ should be equivalent to a single application of the transition matrix constructed at a lag time $k \tau$. We present in \blauw{Fig.~S3} the CK tests for each DNA sequence to demonstrate that the $\tau$ = 1.2 ns models perform very well in predicting transition probabilities out to $k \tau$ = 7.2 ns, validating the Markovian nature and kinetic validity of the four models.

@@

To determine observed rates from the T-jump data, the time-domain t-HDVE data was inverse Laplace transformed into the rate domain using a maximum entropy approach (MEM-iLT).~\citep{Kumar2001OnTimescales} Observed rates were computed from the amplitude-weighted mean rate across detected IR frequency, as previously described.~\citep{Sanstead2018DirectDehybridization} The fast response $k_d^\mathrm{fast}$ is defined purely as this amplitude-weighted mean rate, whereas the dissociation rate constant $k_d^\mathrm{slow}$ was extracted from the observed rate of the slow response $\lambda^\mathrm{slow}$ using a two-state model for self-complementary oligomers,~\citep{Bernasconi2012RelaxationKinetics}
	\begin{equation}\label{e1}
	\lambda^\mathrm{slow} = k_d^\mathrm{slow} + 4[S]_{T_f}k_a^\mathrm{slow},
	\end{equation}
where $[S]_{T_f}$ is the concentration of single-strand oligomer at the final temperature of the T-jump, and $k_a^\mathrm{slow}$ is the association rate constant. \noter{I renamed $k_a$ as $k_a^\mathrm{slow}$ for consistency. Is this OK?} In practice, Eqn.~\ref{e1} is recast in terms of the dissociation equilibrium constant $K_d = k_d^\mathrm{slow} / k_a^\mathrm{slow}$,
	\begin{equation}\label{e2a}
	k_d^\mathrm{slow} = \frac{\lambda^\mathrm{slow}K_d}{K_d+4[S]_{T_f}}.
	\end{equation}
We use this expression to estimate $k_d^\mathrm{slow}$, where we estimate $K_d$ by \rood{...}, $\lambda^\mathrm{slow}$ by \rood{...}, and $[S]_{T_f}$ by \rood{...} \noter{What was written seems incomplete -- it seems like we need to say how we estimate (i) $\lambda^\mathrm{slow}$, (ii) $K_d$, and (iii) $[S]_{T_f}$, not $k_d^\mathrm{slow}$ and $[S]_{T_f}$? Am I understanding Eqn.~2 correctly?} \rood{$k_d^\mathrm{slow}$ and $[S]_{T_f}$ come from melting curves measured with FTIR spectroscopy that are fit to a two-state all-or-nothing dissociation model.~\citep{Marky1987CalculatingCurves} We used the FTIR melting curves previously reported in Ref.~\citep{Sanstead2016}}

@@

We first sought to demonstrate that the coarse-grained 3SPN.2 model accurately recapitulates the experimentally observed kinetics of DNA oligomer hybridization/dissociation by validating our computational predictions against temperature-jump infrared (T-jump IR) experiments. We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted the ``slow'' $k_d^\mathrm{slow}$ (10-30 $\mu$s) and ``fast'' $k_d^\mathrm{fast}$ (70-100 ns) rates estimated from the T-jump response. The slow response has previously been attributed to duplex dissociation on microsecond time scales induced by the rapid heating of the initially hybridized duplex.~\citep{Sanstead2016, Sanstead2018DirectDehybridization} The fast response is has been assigned to terminal base pair fraying,~\citep{Sanstead2016, Sanstead2018DirectDehybridization} with this process corresponding to a relatively complex dynamical process that can span time scales from picoseconds to microseconds.~\cite{Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned Watson-Crick bonds, base-sugar hydrogen bonds, and terminal stacked conformations.~\citep{PinamontiTheModels, Zgarbova2014BaseRNA}

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and fast $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which \rood{both OR either} \noter{Pick the right word.} of the central base pairs first separate to a distance of 1.3 nm. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and defines the dissociated state. We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the Watson-Crick (WC) interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$.

@@

\noter{In the caption to Figure 2: (i) Change panel labels to lower case. (ii) Improve resolution of images. (iii) Label the MSMs in panel c as you did for the bar charts in panel b; AT-all, GC-end, etc. (iv) You nicely label panel a ``Macrostates''. Can you also label the other panels following the labels I have suggested in the caption?}

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{Fig2.pdf}
	\end{center}
\end{figure}
\begin{figure}[ht!]
        \caption{Thermodynamic and kinetic predictions of the sequence-specific MSMs for AT-all, GC-end, GC-core, and GC-mid. (a) \textbf{Macrostates.} Schematic representation of the seven metastable macrostates occupied by one or more of the four sequences: fully hybridized state (H), two-base out-of-register shifted states in the 5$^\prime$ (5S2) or 3$^\prime$ (3S2) direction, four-base out-of-register shifted states in the 5$^\prime$ (5S4) or 3$^\prime$ (3S4) direction, and the fully dissociated state (D). The line drawings represent the 10-base self-complementary sequences, where red-to-blue contacts indicates (possible) WC base pairing and black indicates an unbound bases. Adjacent to each line drawing we provide representative molecular structures corresponding to that macrostate. (b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the \rood{1 ms} \noter{Right?} simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. (c) \textbf{Kinetic networks.} MSMs illustrating the kinetic network for each sequence. The orange circles correspond to the macrostates occupied by each sequence and are labeled by the macrostate codes reported in panel a. The radii of the circles is proportional to the logarithm of the equilibrium occupancy probabilities reported in panel b. Molecular renderings of an illustrative snapshot from the coarse-grained molecular simulations are provided next to each macrostate. The grey arrows between macrostates indicate the presence of a kinetic flux between this pair of states at equilibrium and the arrow thickness is proportional to the flux. (To avoid congesting the diagram, arrows are not reported for fluxes lower than $3\times10^{-6}$ \noter{Units?}.) The numerical value overlaid on each arrow reports the conditional probability that a system occupying the macrostate at the start of the arrow at time $t$ will transition to the macrostate at the end of the arrow by time $(t+\tau)$, where $\tau$=1.2 ns is the lag time corresponding to a single time step of the MSM. Large orange circles correspond to thermodynamically favorable states and large grey arrows correspond to kinetically favorable transitions. (d) \textbf{Kinetic time scales.} Distribution of MSM implied time scales for each sequence. The leading implied time scale corresponds to the characteristic time scale for hybridization/dehybridization and is approximately equal for all systems since simulations were conducted at the same concentration and at the respective melting temperatures. The higher order implied time scales correspond to a spectrum of kinetic relaxations between the constituent macrostates in the MSM corresponding to shifted and/or frayed states.
    }
    \label{fig:allseq_table}
\end{figure}

@@

As illustrated in Fig.~\ref{fig:NN_table}, we see that the NN model tends to predict lower free energies for each macrostate compared to the MSM and therefore overestimate macrostate stability relative to the hybridized (H) state by as much as 14.7 kJ/mol. (The free energy of the hybridized state is set to zero by construction for both the MSM and NN models.) The only exception to this trend is the frayed state (F4) unique to CG-core, where the MSM and NN predictions are in good agreement. Although the trends are in qualitative agreement, what is the root of the quantitative discrepancy of the MSM and NN models in the 5S2, 3S2, 5S4, and 3S4 states? First, the MSM is constructed bottom-up from molecularly detailed 3SPN.2 simulations whereas the NN model is fitted top-down by regression against experimental data. There are approximations inherent in the 3SPN.2 model and the NN model was fitted to limited experimental data assuming a low-order expansion in terms of nearest neighbor additive contributions. Second, although 3SPN.2 is expected to capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the model was not parameterized to fully capture the enthalpic contributions of the interaction of unbound bases with terminal base pairs, whereas this term is explicitly included within the NN model. \noter{Agree with this sentence?} Third, a well known deficiency of the NN model is the absence of any treatment of inert tails -- free bases that extend beyond the dangling end are known to destabilize the duplex.\citep{Michele2014EHybridization}

%Visualizations show that, although free A:T ends in the F4 state could be considered inert tails, they tend to extend away from the intact section of the duplex and therefore may have less of an impact on duplex stability. \noteb{not sure the best way to phrase this, but they could technically still be considered inert tails}. In this case the NN and MSM free energy predictions are in good agreement.

@@

%\noteb{Wrote up a different approach to this section should we present the NN data with no dangling ends instead. Have not fully pieced it together, but this is the basis idea:}

%\bood{As illustrated in Fig.~\ref{fig:NN_table_no_de}, we see that our macrostate free energies recapitulate NN model predictions reasonably well. It is important to note that we do not include dangling end contributions when calculating NN free energies. Although 3SPN.2 can capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the CG model was not parameterized to fully capture the enthalpic contributions of free bases interaction with terminal base pairs. For each macrostate, dangling ends would be expected to have stabilizing effects \noteb{Can show NN with dangling end contributions (original figure) in SI)} and therefore increase state populations compared to what we observe in simulation.}

@@

%\noteb{Can add something here (or in kinetic section) about how, given the MSMs seem not to account for dangling end stability, there would likely be a higher population of each state than we observe in experiment. This could (i) increase out-of-register role in the hybridization process my increasing membership in intermediate states or (ii) decrease the likelihood of shifting behavior due to higher barriers between those states.}

@@

%\begin{figure}[ht!]
%	\begin{center} 
%        \includegraphics[width=160mm, scale=1]{Fig3_no_DE.pdf}
%        \caption{Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TM} The arbitrary additive constant to the free energy is specified such that the free energy of the hybridized macrostate is zero. These NN estimates do not include additional thermodynamic stabilization from dangling ends \noteb{can include this in SI?} (a) Free energy estimates relative to the hybridized state for both the simulated macrostates and corresponding NN prediction. (b) Free energy difference between the NN prediction and simulated value for each macrostate. \noteb{Should the y-axis be more descriptive here?}}
%        \label{fig:NN_table_no_de}
%	\end{center}
%\end{figure}

@@

\section*{TOC Image}

\begin{figure}[ht!]
	\begin{center} 
        \includegraphics[width=150mm, scale=1]{TOC.pdf}
        %\caption*{TOC image}
        \label{fig:TOC}
	\end{center}
\end{figure}

>>>>

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by ultra-fast temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) model that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSummary, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

@@

\textbf{Dimensionality reduction.} The featurized trajectories were then projected into a low-dimensional space in preparation for microstate clustering. The standard approach to doing so is to employ time-lagged independent components analysis (tICA) to learn a linear projection into a low-dimensional embedding that maximally preserves the kinetic variance in the data.~\citep{Perez-Hernandez2013IdentificationConstruction, Schwantes2013ImprovementsNTL9, Noe2015KineticSimulation} In this work, we instead employ state-free reversible VAMPnets (SRVs) that can be conceived of as a nonlinear version of tICA. \citep{Chen} SRVs employ neural networks to learn flexible nonlinear functions of the trajectory featurization that better approximate the slow dynamical modes of the system and have been shown to produce substantially higher resolution MSMs than those developed using tICA in applications to WW-domain and Trp-cage mini-proteins.\citep{Chen, Sidky} Importantly, the SRV modes were learned independently for each system to best approximate the slow collective modes for that particular DNA sequence. SRVs were trained using the SRV package we previously developed (\url{https://github.com/hsidky/srv}) employing the default network architecture of two hidden layers each comprising 100 neurons and tanh activation functions, a learning rate of 0.001, and a batch size of 50,000. We adopted a lag time of $\tau$ = 1.2 ns as appropriately short time scale to resolve the dynamical details of the hybridization/dehybridization dynamics. \citep{Phys2018MarkovValidation} As observed by Husic and Pande, it is conceptually incorrect to treat the lag time as a model hyperparameter to be self-consistently optimized via the VAMP-2 score, but rather as a physically-motivated choice designed to expose the dynamical motions relevant at a particular time and length scale of interest. \citep{Husic2017Note:Selection} As we shall show, this choice of lag time leads to converged and Markovian macrostate MSMs. We guarded against overfitting using five-fold cross-validation in which we divided the trajectory data into 50:50 training:validation splits. We observed plateau of the validation loss and no evidence of overfitting after 20 epochs of training for each system that required approximately 22 GPU-minutes on a single NVIDIA Tesla K80 GPU card. A VAMP-2 scoring of the cumulative kinetic variance explained as a function of number of SRV collective modes retained also showed no evidence of overfitting -- as would be evinced by separation of the training and validation VAMP-2 scores \citep{Sidky} -- and exhibited a knee for each of the four DNA sequences after the fifth, fourth, third, and second slow modes for AT-all, GC-end, GC-core, and GC-mix, respectively (\blauw{Fig.~S1} in the \blauw{Supporting Information}), and motivating the construction of 5D, 4D, 3D, and 2D embeddings, respectively.

@@

\textbf{Macrostate clustering.} The 200 microstates comprising each system were finally coarsened into our terminal macrostate MSM. The microstate transition matrix for each system was computed at a range of lag times $\tau$ and then diagonalized to recover the corresponding eigenvalues $\lambda_i$ and associated implied time scales $t_i = -\tau / \ln \left| \lambda_i \right|$.~\citep{Wehmeyer2019IntroductionV1.0} The implied time scale plots for the four DNA sequences are presented in \blauw{Fig.~S2}. We observe rapid convergence of the implied time scales $t_i$ with lag time $\tau$ for all systems, motivating the construction of high resolution macrostate MSMs at a lag time $\tau$ = 1.2 ns. For this choice of lag time, we recover 5, 4, 2, and 1 implied time scales for the AT-all, GC-end, GC-core, and GC-mix systems, respectively. The identification of $(i-1)$ implied time scales implies the presence of $(i-1)$ slow modes and motivates the coarsening of the system into $i$ macrostates. We estimate these $i$ macrostates by applying PCCA+ spectral clustering to the leading $(i-1)$ eigenvectors of the microstate transition matrix.~\citep{Roblitz2013FuzzyClassification, Weber2018ImplicationsSimulation, Kube2007AConformations} We then estimate the corresponding 6, 5, 3, and 2 macrostate transition matrices $\mathbf{P}$ for the AT-all, GC-end, GC-core, and GC-mix systems, respectively, by projecting the MD simulation trajectories into these discrete macrostates. These macrostate MSMs constitute our terminal kinetic models. We validate the Markovian nature of the four MSMs by subjecting them to the Chapman-Kolmogorov (CK) test \citep{Noe2009ConstructingSimulations,Phys2011MarkovValidation,  Wehmeyer2019IntroductionV1.0}. This test asserts that the transition matrix for a Markovian (i.e., memoryless) MSM constructed at a lag time $\tau$ should satisfy the condition $\mathbf{P}(k \tau) = \mathbf{P}^k(\tau)$, which states that $k$ successive applications of the transition matrix constructed at a lag time $\tau$ should be equivalent to a single application of the transition matrix constructed at a lag time $k \tau$. We present in \blauw{Fig.~S3} the CK tests for each DNA sequence to demonstrate that the $\tau$ = 1.2 ns models perform very well in predicting transition probabilities out to $k \tau$ = 7.2 ns, validating the Markovian nature and kinetic validity of the four models.

@@

To determine observed rates from the T-jump data, the time-domain t-HDVE data was inverse Laplace transformed into the rate domain using a maximum entropy approach (MEM-iLT).~\citep{Kumar2001OnTimescales} Observed rates ($\lambda^\mathrm{fast}$ and $\lambda^\mathrm{slow}$) were computed from the amplitude-weighted mean rate across detected IR frequency, as previously described.~\citep{Sanstead2018DirectDehybridization} The fast response $k_d^\mathrm{fast}$ is defined purely as this amplitude-weighted mean rate, whereas the dissociation rate constant $k_d^\mathrm{slow}$ was extracted from the observed rate of the slow response $\lambda^\mathrm{slow}$ using a two-state model for self-complementary oligomers,~\citep{Bernasconi2012RelaxationKinetics}
	\begin{equation}\label{e1}
	\lambda^\mathrm{slow} = k_d^\mathrm{slow} + 4[S]_{T_f}k_a^\mathrm{slow}
	\end{equation}
where $[S]_{T_f}$ is the concentration of single-strand oligomer at the final temperature of the T-jump, and $k_a^\mathrm{slow}$ is the association rate constant. In practice, Eqn.~\ref{e1} is recast in terms of the dissociation equilibrium constant $K_d$ to solve for $k_d$ and $k_a$.
	\begin{equation}\label{e2a}
	k_d^\mathrm{slow} = \frac{\lambda^\mathrm{slow}K_d(T_f)}{K_d(T_f)+4[S]_{T_f}}
	\end{equation}
	\begin{equation}\label{e2b}
	k_a^\mathrm{slow} = \frac{k_d^\mathrm{slow}}{K_d(T_f)}
	\end{equation}

FTIR temperature series were measured for each sequence as reported previously,~\citep{Sanstead2016} and the second SVD component along temperature was fit to a two-state all-or-nothing model to determine the fraction of intact duplex ($\theta$) as a function of temperature.~\citep{Marky1987CalculatingCurves} $K_d$ and $[S]_{T_f}$ were then determined from $\theta$ at the final temperature of each T-jump measurement. 
	\begin{equation}\label{e3a}
	[S]_{T_f} = [1 - \theta(T_f)]c_{tot}
	\end{equation}
	\begin{equation}\label{e3b}
	K_d(T_f) = \frac{2c_{tot}[1 - \theta(T_f)]^2}{\theta(T_f)}
	\end{equation}
$c_{tot}$ is the total oligonucleotide concentration, which was 2 mM for all measurements.

@@

We first sought to demonstrate that the coarse-grained 3SPN.2 model accurately recapitulates the experimentally observed kinetics of DNA oligomer hybridization/dissociation by validating our computational predictions against temperature-jump infrared (T-jump IR) experiments. We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted the ``slow'' $k_d^\mathrm{slow}$ (10-30 $\mu$s) and ``fast'' $k_d^\mathrm{fast}$ (70-100 ns) rates estimated from the T-jump response. The slow response has previously been attributed to duplex dissociation on microsecond time scales induced by the rapid heating of the initially hybridized duplex.~\citep{Sanstead2016, Sanstead2018DirectDehybridization} The fast response has been assigned to terminal base pair fraying,~\citep{Sanstead2016, Sanstead2018DirectDehybridization} with this process corresponding to a relatively complex dynamical process that can span time scales from picoseconds to microseconds.~\cite{Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix, Galindo-Murillo2015ConvergenceDGCACGAACGAACGAACGC} All-atom simulations suggest that frayed ends can assume misaligned Watson-Crick bonds, base-sugar hydrogen bonds, and terminal stacked conformations.~\citep{PinamontiTheModels, Zgarbova2014BaseRNA}

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and fast $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and defines the dissociated state. We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the Watson-Crick (WC) interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$.

@@

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{Fig2.pdf}
	\end{center}
\end{figure}
\begin{figure}[ht!]
        \caption{Thermodynamic and kinetic predictions of the sequence-specific MSMs for AT-all, GC-end, GC-core, and GC-mix. (a) \textbf{Macrostates.} Schematic representation of the seven metastable macrostates occupied by one or more of the four sequences: fully hybridized state (H), two-base out-of-register shifted states in the 5$^\prime$ (5S2) or 3$^\prime$ (3S2) direction, four-base out-of-register shifted states in the 5$^\prime$ (5S4) or 3$^\prime$ (3S4) direction, and the fully dissociated state (D). The line drawings represent the 10-base self-complementary sequences, where red-to-blue contacts indicates (possible) WC base pairing and black indicates an unbound bases. Adjacent to each line drawing we provide representative molecular structures corresponding to that macrostate. (b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the 1 ms of simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. (c) \textbf{Kinetic networks.} MSMs illustrating the kinetic network for each sequence. The orange circles correspond to the macrostates occupied by each sequence and are labeled by the macrostate codes reported in panel a. The radii of the circles is proportional to the logarithm of the equilibrium occupancy probabilities reported in panel b. Molecular renderings of an illustrative snapshot from the coarse-grained molecular simulations are provided next to each macrostate. The grey arrows between macrostates indicate the presence of a kinetic flux between this pair of states at equilibrium and the arrow thickness is proportional to the flux. (To avoid congesting the diagram, arrows are not reported for fluxes lower than $3\times10^{-6}$ \noteb{flux is in terms of $p_i*p_{ij}$ so I think it should still be unitless?}.) The numerical value overlaid on each arrow reports the conditional probability that a system occupying the macrostate at the start of the arrow at time $t$ will transition to the macrostate at the end of the arrow by time $(t+\tau)$, where $\tau$=1.2 ns is the lag time corresponding to a single time step of the MSM. Large orange circles correspond to thermodynamically favorable states and large grey arrows correspond to kinetically favorable transitions. (d) \textbf{Kinetic time scales.} Distribution of MSM implied time scales for each sequence. The leading implied time scale corresponds to the characteristic time scale for hybridization/dehybridization and is approximately equal for all systems since simulations were conducted at the same concentration and at the respective melting temperatures. The higher order implied time scales correspond to a spectrum of kinetic relaxations between the constituent macrostates in the MSM corresponding to shifted and/or frayed states.
    }
    \label{fig:allseq_table}
\end{figure}

@@

As illustrated in Fig.~\ref{fig:NN_table}, we see that the NN model tends to predict lower free energies for each macrostate compared to the MSM and therefore overestimate macrostate stability relative to the hybridized (H) state by as much as 14.7 kJ/mol. (The free energy of the hybridized state is set to zero by construction for both the MSM and NN models.) The only exception to this trend is the frayed state (F4) unique to CG-core, where the MSM and NN predictions are in good agreement. Although the trends are in qualitative agreement, what is the root of the quantitative discrepancy of the MSM and NN models in the 5S2, 3S2, 5S4, and 3S4 states? First, the MSM is constructed bottom-up from molecularly detailed 3SPN.2 simulations whereas the NN model is fitted top-down by regression against experimental data. There are approximations inherent in the 3SPN.2 model and the NN model was fitted to limited experimental data assuming a low-order expansion in terms of nearest neighbor additive contributions. Second, although 3SPN.2 is expected to capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the model was not parameterized to fully capture the enthalpic contributions of the interaction of unbound bases with terminal base pairs, whereas this term is explicitly included within the NN model. Third, a well known deficiency of the NN model is the absence of any treatment of inert tails -- free bases that extend beyond the dangling end tend to destabilize the duplex.\citep{Michele2014EHybridization}

%Visualizations show that, although free A:T ends in the F4 state could be considered inert tails, they tend to extend away from the intact section of the duplex and therefore may have less of an impact on duplex stability. In this case the NN and MSM free energy predictions are in good agreement.

@@

%\bood{As illustrated in Fig.~\ref{fig:NN_table_no_de}, we see that our macrostate free energies recapitulate NN model predictions reasonably well. It is important to note that we do not include dangling end contributions when calculating NN free energies. Although 3SPN.2 can capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the CG model was not parameterized to fully capture the enthalpic contributions of free bases interaction with terminal base pairs. For each macrostate, dangling ends would be expected to have stabilizing effects (Can show NN with dangling end contributions in SI and therefore increase state populations compared to what we observe in simulation.)

@@

%Can add something here (or in kinetic section) about how, given the MSMs seem not to account for dangling end stability, there would likely be a higher population of each state than we observe in experiment. This could (i) increase out-of-register role in the hybridization process my increasing membership in intermediate states or (ii) decrease the likelihood of shifting behavior due to higher barriers between those states.

@@

%\begin{figure}[ht!]
%	\begin{center} 
%        \includegraphics[width=160mm, scale=1]{Fig3_no_DE.pdf}
%        \caption{Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TM} The arbitrary additive constant to the free energy is specified such that the free energy of the hybridized macrostate is zero. These NN estimates do not include additional thermodynamic stabilization from dangling ends  (a) Free energy estimates relative to the hybridized state for both the simulated macrostates and corresponding NN prediction. (b) Free energy difference between the NN prediction and simulated value for each macrostate.
%        \label{fig:NN_table_no_de}
%	\end{center}
%\end{figure}

@@

\noteb{check out the options for TOC}
\section*{TOC Image}

\begin{figure}[ht!]
	\begin{center} 
        \includegraphics[width=150mm, scale=1]{TOC_alternate2.pdf}
        %\caption*{TOC image}
        \label{fig:TOC}
	\end{center}
\end{figure}
